Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)

### 2021-09-23 Prelim/prelim.Rmd

@@

```{r out.width="90%"}
knitr::include_graphics("fs_graphic.png")
```

@@

- $Z_j = \begin{cases} 
	i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
	\end{cases}$

@@

- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$

- Reduce data to sufficient statistics
  - $H_p = \sum_{i,j} \mathbf{1}_{(i, j) \in h_p}$
  - $H_{p_j} = \sum_{i} \mathbf{1}_{(i, j) \in h_p}$

@@

- **Storage Efficient Indexing (SEI)** - Remove highly unlikely record labels from memory. Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$

>>>>

\begin{columns}
\begin{column}{0.5\textwidth}

\begin{itemize}
\item Record linkage is the task of identifying duplicate records over noisy datasets.

\item Easy with unique identifiers, difficult when faced with errors

\item Far ranging applications in business, public health, and human rights

\end{itemize}

\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
    \begin{center}
     picture
     \end{center}
\end{column}
\end{columns}

@@

```{r out.width="100%"}
knitr::include_graphics("fs_graphic.png")
```

@@

- $Z_j = \begin{cases} 
	i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
	n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
	\end{cases}$

@@

- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$.

- Reduce data to sufficient statistics
  - $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
  - $H_{p_j} = ||r_{p_j}||$
  - $H_p = \sum_{j} H_{p_j}$

@@

- **Storage Efficient Indexing (SEI)** - Store at most small number $R$ many record labels in each $r_{p_j}$, remove highly unlikely record labels from memory. Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$.

- Hashing plus SEI can reduce memory requirements by >99%.
  - Simulation of $20,000 \times 20,000$ linkage problem with 4 fields. Naive approach requires 6.4GB of storage, hashing and SEI requires 90MB.
